---
title: "Support Vector Machines and Neural Networks"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

***Purpose of this Project:***

*This project involves building SVM and neural network regression models to answer a number of questions. We will use the Credit dataset that is part of the ISLR package.*

**Loading required libraries:**
```{r}
library(ISLR)
library(dplyr)
library(glmnet)
library(caret)
```

**Selecting required features:**
```{r}

library(dplyr)
credit_Filtered = Credit[, c("Income", "Limit", "Rating", "Cards", "Age", "Balance")]
summary(credit_Filtered)


```


**Building a linear SVM regression model to predict Sales based on all other attributes ("Income", "Limit", "Rating", "Cards", "Age", "Balance")):**
```{r}

set.seed(1203)
SVM_Model<- train(Income~.,data=credit_Filtered,method="svmLinear",preProcess=c("center","scale"),tuneLength = 10)
SVM_Model

```
*R square is 80.08%*

**Customizing the search grid by checking the model’s performance for C parameter of 0.1,.5,1 and 10 using 2 repeats of 5-fold cross validation. **

```{r}

set.seed(1203)

grid = expand.grid(C= c(0.1,0.5,1,10))

trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

SVM_Model2 <- train(Income ~., data = credit_Filtered, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"), tuneGrid = grid,
tuneLength = 10)

SVM_Model2
```
**Training a neural network model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education"). **
```{r}
Normalization <- preProcess(credit_Filtered[,1:4],method = c("center","scale"))
Norm_data<-predict(Normalization,credit_Filtered)

NNET_Model<- train(Income~.,data=Norm_data,method="nnet",linout=TRUE, trace = FALSE)
NNET_Model
```
*R-square value with best hyperparameters( size=1) is 37.89.*

**Consider the following input: (Limit = 9, Age = 76, Income = 110, Rating = 100, Cards = 1,  Balance = 100). Estimating Sales for this record using the above neuralnet model? **
```{r}
Input <- data.frame(Limit = 9, Age = 76, Income = 110, Rating = 100, Cards = 1,  Balance = 100)

Prediction<- predict(NNET_Model,Input )
Prediction
```

*Estimated Income: 4.669901*
